\documentclass[10pt,french]{letter}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[french]{babel}
\usepackage{a4}
\usepackage{epsfig}
\usepackage{amssymb}
\usepackage{amsmath}

\usepackage{fancyhdr}
\usepackage{hyperref}
\usepackage[left=3cm,right=3cm,top=3cm,bottom=3cm]{geometry}
\usepackage[breakable, skins]{tcolorbox}
\usepackage{enumitem}

% ── Reviewer section header ───────────────────────────────────────────────────
\newcounter{remarkcount}

\newcommand{\reviewer}[1]{%
  \setcounter{remarkcount}{1}%
  \medskip
  \noindent\textbf{\large Reviewer #1}
  \smallskip\par
  \noindent\rule{\linewidth}{0.6pt}
  \medskip
}

% ── Remark box (gray) — numérotation automatique, reset par \reviewer ─────────
\newtcolorbox{remark}{
  breakable, enhanced,
  before={\stepcounter{remarkcount}},
  colback  = gray!12,
  colframe = gray!55,
  fontupper= \small\itshape,
  title    = {\small\bfseries\upshape Remark~\theremarkcount},
  attach boxed title to top left={yshift=-2mm, xshift=4mm},
  boxed title style={colback=gray!55, colframe=gray!55},
  left=3pt, right=3pt, top=5pt, bottom=3pt,
}

% ── Response box (teal left bar) ─────────────────────────────────────────────
\newtcolorbox{response}{
  breakable, enhanced,
  colback  = teal!5,
  colframe = teal!60,
  leftrule = 3pt, rightrule=0.4pt, toprule=0.4pt, bottomrule=0.4pt,
  fontupper= \small,
  title    = {\small\bfseries Response},
  attach boxed title to top left={yshift=-2mm, xshift=4mm},
  boxed title style={colback=teal!60, colframe=teal!60},
  left=6pt, right=3pt, top=5pt, bottom=3pt,
}

\pagestyle{fancy}

\cfoot{\tiny Laboratoire de Mathématiques, UMR CNRS 5127 --- UFR SceM, campus scientifique --- 73376 Le-Bourget-du-Lac Cedex }

\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}


\def\NAME{Romain Negro}
\address{\NAME \\
Laboratoire de Mathématiques, UMR CNRS 5127\\
UFR SceM, Campus scientifique \\
73376 Le-Bourget-du-Lac Cedex\\
}
\signature{\NAME}
\date{Le-Bourget-du-Lac, le \today}

\pagestyle{fancy}

\begin{document}

  \mbox{~}\hfill\mbox{Le-Bourget-du-Lac, April 25, 2026}
  \medskip

  Dear editors,

  Please find with the letter the revised paper (one with the changes highlighted, one without).

  We would like to thank the reviewers for their thorough reading of our manuscript and their
  constructive remarks, which helped us improve the quality of the paper significantly.
  We have carefully addressed all the comments and concerns raised, and provide below a
  detailed account of the changes made in response to each remark.

  \reviewer{1}

  \begin{remark}
    Usually the visibility problem defines two points as visible if and only if the straight-line
    segment joining them lies completely in the free (empty) space. Your definition of visibility
    (page 1, right column, bottom) defines two points as visible if such line segment lies close
    enough to set X. This seems contradictory with the usual definition (and with the intuition)
    unless X is the set of empty voxels (but the reader expects that it is the set of occupied voxels)
    or you address a different problem wrt the classic one. In any case, this needs to be clarified.
    If the addressed problem is not clear, the overall paper cannot be understood properly.
  \end{remark}
  \begin{response}
    You can picture the classic definition of visibility applied to our work as follows: instead
    of considering the surface of the object as the surface of a volume, we consider the surface
    as a thin layer of voxels (like a shell) and we ask for the visibility to be defined as the
    existence of a straight-line segment joining two points that lies completely in this thin layer.
    We added a figure to better illustrate this point (Figure~3 in the revised version) and linked
    it to the definition of visibility in the paper. We hope this will clarify the definition of
    visibility we are using in our work.
  \end{response}

  \begin{remark}
    The second paragraph on page 2, right column, cannot be understood in this place, as the reader
    does not know how the proposed method works. It can be maybe better moved to the conclusions.
  \end{remark}
  \begin{response}
    We moved this paragraph to the conclusion section, as suggested by the reviewer.
  \end{response}

  \begin{remark}
    The paper needs a table of used symbols. Many symbols are used once, without saying that in the rest of
    the paper such symbol will always denote such entity, and then used several lines or pages later, where
    the reader has forgotten what they denote. Moreover, the use of symbols is not always consistent.
  \end{remark}
  \begin{remark}
    The digital set is sometimes denoted with X and other times with Z (for example on page 3, left column,
    both symbols are used). I suggest not to use Z because it is too similar to the set of integer numbers.
    Symbol K seems to be used with the same meaning as X, Z in the right column of page 3.
  \end{remark}
  \begin{remark}
    Symbol h is used on page 2 (right column, top), on page 4 (right column, bottom) and only on page 6 it
    is defined as the gridstep (not in main text, but within the caption of fig.6).
  \end{remark}
  \begin{response}
    Considering how many symbols are used in the paper, we think that a table of symbols would be too heavy
    and would not be really helpful for the reader. However, we added a notation section at the start of the
    visibility section, where we introduce the main symbols used in the paper and explain what they denote.
    We also made sure to be consistent with the use of symbols throughout the paper.
  \end{response}

  \begin{remark}
    The figures, connecting the points of the digital set with bold line segments (like figs. 1 and 2),
    are misleading. They lead to think that X is a 1D complex consisting of points and edges, whereas
    it is a set of points.
  \end{remark}
  \begin{remark}
    The second paragraph of the left column of page 4 seems contraddictory. It says that the input for
    visibility is a subcomplex, while previously it was a set of points in digital space.
  \end{remark}
  \begin{response}
    There was indeed an inconsistency in the figures: the bold line segments connecting lattice
    points made the digital set look like a 1D complex of points and edges, rather than a plain set
    of points. We removed those segments from Figures 1 and 2, and now display only the lattice
    points together with the 1-cells of the cubical complex to make the distinction clear.

    Regarding the apparent contradiction raised in Remark~\the\numexpr\value{remarkcount}-1\relax:
    there is no inconsistency between "a set of points in digital space" and "a subcomplex". The
    key is the lattice map (Khalimsky encoding), which assigns to every cell of the cubical
    complex $\mathcal{C}^d$ a unique point with integer coordinates in $\mathbb{Z}^d$. Through this
    bijection, any cell subcomplex (in particular the $k$-star $\mathrm{Star}_k(Z)$ of the digital
    set $Z$) can equivalently be described as a set of points in $\mathbb{Z}^d$.
  \end{response}


  \begin{remark}
    The lattice map (page 4) must be explained better. Figure 3 does not help, unless it is
    explained in more detail in the text. The caption talks about 63 cells, but I cannot
    understand what cells are they in the figure. The process of passing from the cell complex
    to the intervals must be explained.
  \end{remark}
  \begin{response}
    We took inspiration from the original paper's DGMM presentation of the lattice map, and added
    a more detailed explanation of how the lattice map works and how it allows to pass from the cell
    complex to the intervals' representation. We hope this will clarify the concept of lattice map
    and how it is used in our work.
  \end{response}

  \begin{remark}
    Figure 4 (the working example of the algorithm) must be explained in the text, otherwise it is
    useless. What are Vi , Fi , Ti and Ri appearing in the figure? The algorithm is one of the main
    contributions of the paper, so the reader must be enabled to understand it.
  \end{remark}
  \begin{response}
    We added a much more detailed explanation of the working example of the algorithm, including a
    step-by-step description of what is happening in the figure and what the different sets Vi, Fi,
    Ti and Ri represent.
  \end{response}

  \begin{remark}
    Figure 5: the source point (now blue on grey) should be marked in another, more evident, color,
    for example red. Does this figure show 1-visibility?
  \end{remark}
  \begin{response}
    We redesigned Figure~5 (Figure~6 in the revised version) to make the source points more visible
    (now in red and slightly bigger), and set the visible points in yellow on newer images. We also
    extended the visibility radius to its maximum value to avoid any confusion about what visibility
    is being shown in the figure. The figure do show the 1-visibility as noted in the visibility
    introductory text ("Unless otherwise specified, we will consider visibility with k = 1").
  \end{response}

  \begin{remark}
    Page 7, top of left column: the visibility normal is the last eigenvector... The normal should be
    a unit vector. Do you mean that it is the eigenvector is taken after normalization, or do you
    guarantee that it 1 has norm = 1? Please clarify.
  \end{remark}
  \begin{response}
    By definition, the eigenvectors of a covariance matrix are unit vectors, so there is no need to
    normalize them, nor to precise that they are unit vectors.
  \end{response}

  \begin{remark}
    Page 7, top of right column: convergence is proven starting from k = 2. Do you mean for any
    k $\geq$ 2 or for any k $\leq$ 2?
  \end{remark}
  \begin{response}
    We meant for any k $\geq$ 2, as stated in the convergence theorem (Section 4) and implied
    by the wording. We clarified this point in the text using the notation k $\geq$ 2 instead of k = 2.
  \end{response}

  \begin{remark}
    Figure 13: I note that the first three rows and columns present vertical artifacts on some faces,
    while in the rightmost column (new method) they don’t appear. I think it is worth to investigate
    why. It is a good feature of the proposed method to avoid such artifacts!
  \end{remark}
  \begin{response}
    Those vertical artifacts are not the only artifacts present in the first three rows and columns
    of Figure 13, but they are indeed more visible than the others. We added a paragraph in the
    section discussing the results of Figure 13. Inside, we explain that those artifacts are due to the fact that
    the radii of II and PCA normal estimation methods lack the ability to adapt to the local geometry (either
    being too small and capturing inherent digital noise, or being too large and smoothing out salient features).
    However, our method is able to adapt to the local geometry and thus avoid such artifacts. We also added two
    versions of radii for these methods in the figure to illustrate this point: one with a small radius (above
    the diagonal) and one with a large radius (below the diagonal).
  \end{response}

  \begin{remark}
    Page 10, top of right column. You should explain what does it mean that the boundary has
    positive reach (what is the reach?). Here X denotes a compact subset of the real space, it
    denoted a digital set of points elsewhere.
  \end{remark}
  \begin{response}
    The reach of a set $\partial X$ (introduced by Federer, 1959) is the largest $\rho > 0$
    such that every point within distance $\rho$ of $\partial X$ has a unique nearest point on
    $\partial X$. Intuitively, it measures how "round" the boundary is: a sphere of radius $r$
    has reach $r$, while a sharp corner has reach $0$. The condition $\mathrm{reach}(\partial X) > \rho > 0$
    therefore ensures that the boundary is smooth enough (no sharp features) for the convergence
    proof to hold.
    We also clarified the notation issue concerned with $X$ and $Z$ in the paper,
    and made sure to use $X$ consistently to denote the compact subset of $\mathbb{Z}^d$ and $Z$ to
    denote the digital sets.
  \end{response}

  \begin{remark}
    I did not follow the proof because I don’t have the necessary mathematical knowledge. I hope
    that the other reviewers can cover this part. The term ”sandwiched” (page 11, right column,
    line 5 from bottom) should be probably replaced by a more technical one. Page 13, top: what
    does it mean ”at most” orthogonal? In the caption of Fig. 14, ”1st” should be ”first”.
  \end{remark}
  \begin{response}
    The term "sandwiched" sounds indeed a bit informal, but it is a common term in mathematics
    to describe a situation where an object is "trapped" between two other objects. In our case, the
    estimated normal vector is "sandwiched" between the true normal vector and the tangent space of the
    surface, meaning that it is close to both of them. We fixed the several typos you pointed out.
  \end{response}

  \reviewer{2}


  \begin{remark}
    There are many places where it could have been more precise. For instance:
    "there is a straight segment joining them that stays in the close vicinity of
    the surface" - what is the relevance of this close vicinity and the visibility?
  \end{remark}
  \begin{response}
    The relevance of the "close vicinity" is precisely what makes our definition of visibility
    specific to digital surfaces. Our framework considers visibility within a tubular
    neighborhood of the surface: two points $p$ and $q$ are visible if the straight-line segment
    $[p,q]$ stays at $L_\infty$-distance strictly less than $k$ from the digital set $Z$:
    it never leaves the tube of thickness $k$ around $Z$. This is equivalent to asking that the
    segment remains inside the $k$-star $\mathrm{Star}_k(Z)$ of the digital set throughout its
    length. In other words, the "close vicinity of the surface" is the visibility
    condition: requiring the segment to stay near the surface and requiring it to be fully visible
    (in our sense) are one and the same thing.
  \end{response}

  \begin{remark}
    "several days of computation for 1M triangles at the time" - maybe this was
    true for extreme precision computations in the 1990s? With the current GPU, it
    takes only milliseconds. So, it could be an overexaggeration that can be
    removed.
  \end{remark}
  \begin{response}
    The reviewer is right that modern GPU-accelerated ray tracing can answer approximate
    visibility queries (whether a single pair of points is mutually visible) in
    milliseconds. However, the sentence refers to computing the exact, complete visibility
    complex, in other words, determining all mutually visible pairs across the entire surface, which is a
    fundamentally different and much harder problem. For $n$ triangles, exact 3D visibility
    algorithms operate in $O(n^4)$ in the worst case (as recalled in the introduction), leading to
    roughly $10^{24}$ operations for $n = 10^6$ triangles. On a sequential CPU, and even on a
    highly parallel GPU, such complexity does not become tractable simply by virtue of faster
    hardware. We altered the original sentence in the revised introduction to make this distinction
    explicit and avoid any misinterpretation.
  \end{response}

  \begin{remark}
    I like the second paragraph in the introduction. Still, the shift from
    traditional visibility to lattices in the third paragraph should be explained
    more clearly (this includes explaining how lattices are related and what
    changes/challenges they bring). This will make it very easy to understand.
  \end{remark}
  \begin{response}
    We added a more detailed explanation of the shift from traditional visibility
    to lattices in the transition from how the algorithm works to how we exploit
    the lattice map and the intervals representation page 6 of the changes.
  \end{response}

  \begin{remark}
    While efficiency is highlighted in the introduction, the only comparison was
    with BF visibility, and the computation time is almost the same or even higher
    (for radius=30), as the paper rightly notes. Either tone down the emphasis on
    efficiency or add better comparisons.
  \end{remark}
  \begin{response}
    We agree that our method does not outperform the brute-force method in terms of
    raw computation time, but it is important to note that our method computes the
    exact, complete visibility complex, while the brute-force method only
    computes an approximate, partial visibility complex (since it relies on
    its connectedness assumption to reduce the number of pairs of points it checks)
    in the same time complexity. In other words, our method achieves the same efficiency
    as the brute-force method, but for a much stronger notion of visibility. We
    removed the emphasis on efficiency in the introduction to avoid any confusion.
  \end{response}

  \begin{remark}
    I'm not very convinced of the claim on "covariance-based normal estimation
    methods." Isn't it possible to reuse them on the surface? What exactly doesn't
    hold for "digitized shapes" and what are their implications? This is important
    for motivation.
  \end{remark}
  \begin{response}
    The main issue with covariance-based normal estimation methods on digitized shapes is that they rely
    on the noise being a gaussian perturbation of the underlying smooth surface, which is not the case
    for digitization processes. In particular, the digitization process can introduce non-gaussian noise
    and can create sharp features that are not present in the original smooth surface, which can lead to
    inaccurate normal estimation. We can observe this issue in the results of Figure 13 for the PCA normal
    estimation method, where it either captures the inherent digital noise (with the oscillations on
    smaller radii) or smooths out salient features (with larger radii). Our method, on the other hand,
    is designed to be aware of the local geometry and can adapt to it without parameters. It avoids
    above issues while providing accurate normal estimation even in the presence of digitization artifacts.
    We pinpointed this issue in the figure 13 discussion section of the revised version to better motivate
    our method and highlight its advantages over covariance-based normal estimation methods.
  \end{response}

  \begin{remark}
    I believe "Non-Iterative, Feature-Preserving Mesh Smoothing" deserves a
    citation, as their solutions (the core idea for normal estimation) are somewhat
    similar.
  \end{remark}
  \begin{response}
    We added a citation to "Non-Iterative, Feature-Preserving Mesh Smoothing" in the introduction section
    when we discuss the motivation for our method and the related work on normal estimation. We do think
    that this paper is relevant to our work, as it has similar hypotheses than II. However, our method
    is different in that it does not require any parameter tuning and is designed to be
    specifically aware of the local geometry of digitized shapes, which allows it to provide normal estimation
    independent of the scale and without smoothing out salient features, while still being accurate and convergent.
  \end{response}

  \begin{remark}
    I do not fully understand the last paragraph of the introduction. Can you add
    an example in the discussion?
  \end{remark}
  \begin{response}
    Please note that the last paragraph of the introduction has been moved to the conclusion section,
    as suggested by the first reviewer. Here is a simple example to illustrate the point: consider
    a simple triangle in 3D space, and take its decomposition into lattice points. We can use this
    decomposition as the input for our algorithm instead of the decomposition of a segment. The
    algorithm will then return the points on the surface where this triangle can be found, which
    corresponds to the points where this specific triangle can be "placed" on the surface. This
    illustrates how our method can be used to analyze the local geometry of the surface by looking at
    the visibility of specific shapes (in this case, a triangle) on the surface.
  \end{response}

  \begin{remark}
    I like the idea of visibility by the intersections of integer intervals. However, adding
    intuitions/motivations will improve readability.
  \end{remark}
  \begin{response}
    Please refer to remark 8 of the first reviewer.
  \end{response}

  \begin{remark}
    Considering that there are no page limits (?), I would suggest explaining the
    algorithms in detail.
  \end{remark}
  \begin{remark}
    It took me quite some time to decode Figure 4. Maybe it can be improved? Or
    the captions could have been more precise? And maybe annotate by noting exactly
    what each piece says.
  \end{remark}
  \begin{response}
    Please refer to remark 9 of the first reviewer.
  \end{response}

  \begin{remark}
    Does the blue point represent the "source point" in Figure 5? I see a small
    chirp on the blue point, so I thought the blue point lies on the surface? If
    yes, then the visible points don't make sense to me.
  \end{remark}
  \begin{remark}
    It is not clear to me why the "visbility does not cross the gap between the
    two parts" - how does the gap matter for visibility? Or is it because of your
    threshold that you are ignoring far-away, visible points?
  \end{remark}
  \begin{response}
    Please refer to remark 10 of the first reviewer.
    Regarding the last remark (the gap and visibility): our definition of visibility is
    specifically designed to capture visibility along the surface rather than through free
    space. The key consequence is that two points sitting on two nearby but geometrically
    disconnected parts of the surface (for instance, on opposite walls of a thin gap or
    cavity) are not mutually visible in our sense. This is because the straight segment
    joining them would leave the $k$-star of the surface and cross empty space. This is
    not a threshold effect: it is a fundamental property of the definition.

    This behavior is precisely what we exploit for normal estimation. When computing the visibility
    cone at a point $p$, we only collect information from the same connected surface patch; points
    on the other side of a gap or sharp feature are invisible, so their (potentially contradictory)
    normal information cannot pollute the estimate at $p$.

    The maximum visibility radius is a separate, optional parameter: it simply
    caps how far we search for visible points, discarding points that are visible but very distant.
    Since the normal estimate changes very little once enough nearby visible points are collected,
    this truncation trades a negligible loss of accuracy for a significant reduction in computation
    time. We updated Figure~5 (Figure~6 in the revised version) to show visibility pushed to the full extent of
    the shape (no radius cap), making the gap-blocking and feature-blocking behavior clearly visible.
  \end{response}

  \begin{remark}
    I'm not very convinced about the "weighted centroid of the visibile points
    around" sentence. What happens at sharp corners with some weird distribution
    around them?
  \end{remark}
  \begin{response}
    The "weird distributions" around sharp corners are indeed a concern for normal estimation, as they
    can lead to inaccurate estimates if not handled properly. However, our method is designed to be robust
    to such situations. As the gridstep decreases, the distribution of visible points around sharp corners
    becomes more regular and disk-like, which allows the weighted centroid to provide a good estimate of
    the normal direction. Moreover, for points close to salient features, the visibility cone will naturally
    include more points from the non-salient part of the surface, which helps to mitigate the influence
    of the salient feature on the normal estimation. In other words, our method leverages the visibility
    structure to adaptively weight the contribution of visible points, providing accurate normal estimates
    even in the presence of sharp features and irregular distributions.
  \end{response}

  \begin{remark}
    Unfortunately, I feel the experimentation section is too weak. The comparisons
    are made with very old baseline techniques. There are many recent techniques,
    especially the neural ones. Did you check how the proposed method compares to
    them? Actually, this is the second most important concern I have.
  \end{remark}
  \begin{response}
    We focused our comparisons on classical, well-established methods with known theoretical properties
    (in particular, convergence guarantees) to provide a solid baseline for evaluating our method. We did
    not find any recent deep learning-based methods specifically designed to compute normals or curvatures
    from digital surfaces. The closest we found were methods that take point clouds as input to compute
    normals, but the results displayed in the papers were not impressive (worse than II in our context).
    However, after performing extensive tests on Adafit, we realized that its results were much better
    than those proposed in the publications. We therefore decided to add a comparison with Adafit in
    the experimentation section, in addition to the II and PCA methods already present. We also added
    a discussion on the results obtained with Adafit as well as its advantages and disadvantages compared
    to our method. We hope this will strengthen the experimentation section and provide a more comprehensive
    comparison with existing methods.
  \end{response}

  \begin{remark}
    I'm not very convinced about the results shown in Figure 12. I would say the
    PCA is looking much better, at least for the dragon.
  \end{remark}
  \begin{response}
    Figure~12 is now Figure~13 in the revised version. We added a discussion of the results shown in this figure,
    including a comparison between our method and PCA. We acknowledge that PCA seems to perform better in some cases
    (like the tie fighter), but it is important to note that PCA's performance is highly sensitive to the choice of
    radius: with a small radius, it captures the inherent digital noise and produces oscillations, while with a large
    radius, it smooths out salient features and loses important geometric details. Our method, on the other hand, is
    designed to be parameter-free and adapt to the local geometry, providing accurate normal estimation without
    smoothing out features or being affected by noise. We highlighted this advantage of our method in the discussion
    of Figure~13 and by including a comparison with both radii inside this figure to provide a more balanced
    evaluation of the results.
  \end{response}

  We also made several minor improvements to the paper, including fixing typos, improving the clarity of some
  explanations, and enhancing the figures for better readability. We hope this revision and the above answers
  will satisfy all your requests and concerns, and are looking forward to hearing from you soon.

  Best regards
  \bigskip

  \mbox{~}
  \hfill
  \begin{minipage}{0.32\textwidth}
    Romain Negro \\
    PhD in Computer Science\\
    University Savoie Mont Blanc\\
    CNRS, LAMA
  \end{minipage}
  \hfill
  \begin{minipage}{0.32\textwidth}
    Jacques-Olivier Lachaud\\
    Professor in Computer Science\\
    University Savoie Mont Blanc\\
    CNRS, LAMA
  \end{minipage}
  \hfill
  \mbox{~}
\end{document}

